---
title: "test"
author: "Colleen Minnihan"
date: "4/7/2021"
output: html_document
---
```{r libraries, message=FALSE}
# SEE modeldata package for new datasets
library(tidyverse)         # for graphing and data cleaning
library(tidymodels)        # for modeling
library(stacks)            # for stacking models
library(naniar)            # for examining missing values (NAs)
library(lubridate)         # for date manipulation
library(moderndive)        # for King County housing data
library(DALEX)             # for model interpretation  
library(DALEXtra)          # for extension of DALEX
library(patchwork)         # for combining plots nicely
library(dbplyr)            # for SQL query "cheating" - part of tidyverse but needs to be loaded separately
library(mdsr)              # for accessing some databases - goes with Modern Data Science with R textbook
library(RMySQL)            # for accessing MySQL databases
library(RSQLite)           # for accessing SQLite databases

#mapping
library(maps)              # for built-in maps
library(sf)                # for making maps using geom_sf
library(ggthemes)          # Lisa added - I like theme_map() for maps :)

#tidytext
library(tidytext)          # for text analysis, the tidy way!
library(textdata)          
library(reshape2)
library(wordcloud)         # for wordcloud
library(stopwords)

theme_set(theme_minimal()) # Lisa's favorite theme
```


## SQL

You will use the `airlines` data from the SQL database that I used in the example in the [tutorial](https://advanced-ds-in-r.netlify.app/posts/2021-03-29-sqlinr/). Be sure to include the chunk to connect to the database here. And, when you are finished, disconnect. You may need to reconnect throughout as it times out after a while.

```{r}
#connecting to the airlines database
con_air <- dbConnect_scidb("airlines")
```

SQL code chunk! To see variable names
```{sql connection=con_air}
DESCRIBE flights;
```



**Tasks**:

1. Create a SQL chunk and an equivalent R code chunk that does the following: for each airport (with its name, not code), year, and month find the total number of departing flights, the distinct destinations to which they flew, the average length of the flight, the average distance of the flight, and the proportion of flights that arrived more than 20 minutes late. In the R code chunk, write this out to a dataset. (HINT: 1. start small! 2. you may want to do the R part first and use it to "cheat" into the SQL code).

SQL chunk
```{sql connection=con_air}

SELECT * 
FROM flights 
LIMIT 20
```


for each airport (with its name, not code), year, and month find the total number of departing flights, the distinct destinations to which they flew, the average length of the flight, the average distance of the flight, and the proportion of flights that arrived more than 20 minutes late. In the R code chunk, write this out to a dataset. (HINT: 1. start small! 2. you may want to do the R part first and use it to "cheat" into the SQL code).

R chunk
```{r}
dbListTables(con_air)

flights <- 
  tbl(con_air,
      sql("
          SELECT * 
          FROM flights 
          LIMIT 2000
          ")
      )

flights <- flights %>%
  select(year, month, origin, dest, air_time, distance, arr_delay)

each_airport <- flights %>%
  group_by(origin, month) %>%
  summarize(num_depart = count(origin),
            avg_flight_dist = mean(distance),
            prop_late = (arr_delay > 20)/(arr_delay))

each_airport
# for each airport (with its name, not code) and month find the total number of departing flights, the average distance of the flight, and the proportion of flights that arrived more than 20 minutes late.
```

SQL Chunk
```{sql connection = con_air}
SELECT 
  origin, month,
  COUNT(origin) AS num_depart,
  AVG(distance) AS avg_flight_dist,
  AVG(arr_delay > 20) AS prop_late_over20
FROM (SELECT * FROM flights LIMIT 2000) `fl`
GROUP BY carrier, month
```

